---
title: "RF Model"
output: html_notebook
---




# Load data from stats.Rmd
```{r}
load('/Volumes/Arges_NFS/PROJECTS/mcgaughey/EGA_EGAD00001002656_NGS_reanalyze/clean_data/gemini_out_variants_het_and_hom.Rdata')
```

# Set up data for modeling
```{r}
all_PATH <- all %>% 
  select(variant_id, filter, type, call_rate, num_hom_ref, num_het, num_hom_alt, aaf, is_exonic, is_coding, is_lof, is_splicing, impact, impact_so, impact_severity, polyphen_score, sift_score, aaf_1kg_afr_float:an_exac_sas, max_aaf_all, pli, pnull, precessive, phylop_100way, canonical, cadd_phred, Genotype, Status, Complicated_Status) %>% 
  filter(Status=='Pathogenic') %>% 
  unique()
all_NOT_PATH <- all %>% 
  select(variant_id, filter, type, call_rate, num_hom_ref, num_het, num_hom_alt, aaf, is_exonic, is_coding, is_lof, is_splicing, impact, impact_so, impact_severity, polyphen_score, sift_score, aaf_1kg_afr_float:an_exac_sas, max_aaf_all, pli, pnull, precessive, phylop_100way, canonical, cadd_phred, Genotype, Status, Complicated_Status) %>% 
  filter(Status=='NotPathogenic') %>% 
  unique()

# multiply path by 10x with duplications
#all_PATH_10X <- bind_rows(replicate(10, all_PATH, simplify = FALSE))
all_PATH_10X <- all_PATH
# cut down pathogenic to 10x of path
set.seed(115470)
all_NOT_PATH_CUT <- all_NOT_PATH #%>% sample_n(nrow(all_PATH_10X) * 15)

ML_set <- rbind(all_PATH_10X, all_NOT_PATH_CUT)

#cadd phred has NA
ML_set[is.na(ML_set)] <- 0
# convert chr to factor
ML_set <- ML_set %>%  mutate_each_( funs(as.factor(.)), names( .[,sapply(., is.character)] ))
# Fix numeric that are fctr
ML_set$polyphen_score <- as.numeric(ML_set$polyphen_score)
ML_set$sift_score <- as.numeric(ML_set$sift_score)
ML_set$adj_exp_lof <- as.numeric(ML_set$adj_exp_lof)
ML_set$adj_exp_mis <- as.numeric(ML_set$adj_exp_mis)
ML_set$adj_exp_syn <- as.numeric(ML_set$adj_exp_syn)
ML_set$pli <- as.numeric(ML_set$pli)
ML_set$pnull <- as.numeric(ML_set$pnull)
ML_set$precessive <- as.numeric(ML_set$precessive)
ML_set$phylop_100way <- as.numeric(ML_set$phylop_100way)
ML_set$Status <- factor(ML_set$Status, levels=c('Pathogenic','NotPathogenic'))


```

## One hot encoding (Dummy variables)
Turn categorical factors into variables

Many ML algorithms can't handle factor

You can feed this into the `Train and Test sets` section below if you are using something like a GLM or NN. 

Skip for now
```{r, eval = F}
temp <- ML_set %>% dplyr::select(-filter, -Status, -Complicated_Status)
dmy <- dummyVars(" ~ .", data = temp, fullRank=T)
ML_set_dummy <- data.frame(predict(dmy, newdata = temp))
ML_set_dummy <- ML_set_dummy %>% mutate(Status = ML_set$Status, Complicated_Status = ML_set$Complicated_Status)
```

## Train and Test sets
Split in training / testing

```{r}
set.seed(115470)
train_set <- ML_set %>% 
  group_by(Status, Complicated_Status) %>% 
  sample_frac(0.5) %>% ungroup()

test_set <- ML_set %>% 
  filter(!variant_id %in% train_set$variant_id)
```

## Recipe
skip for now
```{r, eval=FALSE}
library(caret)
train_setR <- train_set %>% dplyr::select(-Complicated_Status, -variant_id)
rec_obj <- recipe(Status ~ ., data = train_setR)
```



## RF
This is the only model that kind of works

Have quickly tried keras, glm, nnet - they are disastrously bad. Probably need to balance classes for another approach to work. 
```{r}
library(caret)
library(mlbench)
library(parallel)
library(doParallel)
#library(plotROC)
fitControl <- trainControl(## 5-fold CV
  method = "LOOCV",
  number = 5,
  #classProbs=T,
  #savePredictions = T,
  allowParallel = T)
cluster <- makeCluster(detectCores() - 4) # leave some cores for me!
registerDoParallel(cluster)

set.seed(825)
rfFit <- caret::train(x=train_set %>% dplyr::select(-variant_id, -Status, -Complicated_Status) %>% data.frame(), 
                      y = train_set$Status, 
                      preProcess = c("scale", "center"),
                      method = "rf", 
                      metric = 'Kappa',
                      trControl = fitControl)
rfFit
stopCluster(cluster)
registerDoSEQ() # forces R down to single thread - can't do parallel run after this?


confusionMatrix(data = rfFit$pred$pred, reference = rfFit$pred$obs, mode='prec_recall')
# change cutoff
confusionMatrix(data = rfFit$pred %>% mutate(pred2 = case_when(Pathogenic > 0.2 ~ 'Pathogenic', TRUE ~ 'NotPathogenic')) %>% pull(pred2), reference = rfFit$pred$obs)

randomForest::varImpPlot(rfFit_tissue$finalModel)
selectedIndices <- rfFit$pred$mtry == rfFit_tissue$bestTune[1,] #26

g <- ggplot(rfFit$pred[selectedIndices, ], aes(m=Pathogenic, d=factor(obs, levels=c('Pathogenic','NotPathogenic')))) + 
  geom_roc(n.cuts=0) + 
  coord_equal() +
  style_roc()

g + annotate("text", x=0.75, y=0.25, label=paste("AUC =", round((calc_auc(g))$AUC, 4)))
```