---
title: "RF Model"
output: html_notebook
---




# Load data from stats.Rmd
```{r}
# load('/Volumes/Arges_NFS/PROJECTS/mcgaughey/EGA_EGAD00001002656_NGS_reanalyze/clean_data/gemini_out_variants_het_and_hom.Rdata')
load('/Volumes/eyeMac/Volumes/Arges/PROJECTS/mcgaughey/EGA_EGAD00001002656_NGS_reanalyze/clean_data/gemini_out_variants_het_and_hom.Rdata')
```

# Set up data for modeling
```{r}
library(tidyverse)
all_PATH <- all %>% 
  select(variant_id, filter, type, call_rate, num_hom_ref, num_het, num_hom_alt, aaf, is_exonic, is_coding, is_lof, is_splicing, impact, impact_so, impact_severity, polyphen_score, sift_score, aaf_1kg_afr_float:an_exac_sas, max_aaf_all, pli, pnull, precessive, phylop_100way,  fitcons_float, lof_z, mis_z, syn_z, gerp_elements, grantham, cadd_phred, Genotype, Status, Complicated_Status) %>% 
  filter(Status=='Pathogenic') %>% 
  unique()
all_NOT_PATH <- all %>% 
  select(variant_id, filter, type, call_rate, num_hom_ref, num_het, num_hom_alt, aaf, is_exonic, is_coding, is_lof, is_splicing, impact, impact_so, impact_severity, polyphen_score, sift_score, aaf_1kg_afr_float:an_exac_sas, max_aaf_all, pli, pnull, precessive, phylop_100way,  fitcons_float, lof_z, mis_z, syn_z, gerp_elements, grantham, cadd_phred, Genotype, Status, Complicated_Status) %>% 
  filter(Status=='NotPathogenic') %>% 
  unique()

all_set <- rbind(all_PATH, all_NOT_PATH)
process <- function(df) {
  #cadd phred has NA
  df[is.na(df)] <- 0
  # convert chr to factor
  df <- df %>%  mutate_each_( funs(as.factor(.)), names( .[,sapply(., is.character)] ))
  # Fix numeric that are fctr
  df$polyphen_score <- as.numeric(df$polyphen_score)
  df$sift_score <- as.numeric(df$sift_score)
  df$adj_exp_lof <- as.numeric(df$adj_exp_lof)
  df$adj_exp_mis <- as.numeric(df$adj_exp_mis)
  df$adj_exp_syn <- as.numeric(df$adj_exp_syn)
  df$pli <- as.numeric(df$pli)
  df$pnull <- as.numeric(df$pnull)
  df$precessive <- as.numeric(df$precessive)
  df$phylop_100way <- as.numeric(df$phylop_100way)
  df$fitcons_float <- as.numeric(df$fitcons_float)
  df$lof_z <- as.numeric(df$lof_z)
  df$mis_z <- as.numeric(df$mis_z)
  df$syn_z <- as.numeric(df$syn_z)
  df$gerp_elements <- as.numeric(df$gerp_elements)
  df$grantham <- as.numeric(df$grantham)
  df$Status <- factor(df$Status, levels=c('Pathogenic','NotPathogenic'))
  df
}

all_set <- process(all_set)


all_PATH <- all_set %>% filter(Status=='Pathogenic')
# cut down pathogenic to 5x of path
set.seed(115470)
all_NOT_PATH__CUT <- all_set %>% filter(Status=='NotPathogenic', max_aaf_all < 0.02) %>% sample_n(nrow(all_PATH_10X) * 5)

ML_set <- rbind(all_PATH, all_NOT_PATH__CUT)



```

## One hot encoding (Dummy variables)
Turn categorical factors into variables

Many ML algorithms can't handle factor

You can feed this into the `Train and Test sets` section below if you are using something like a GLM or NN. 

```{r, eval = F}
library(dummies)
temp <- ML_set %>% dplyr::select(-filter, -Status, -Complicated_Status)
temp <- dummy.data.frame(temp, sep='_')
ML_set_dummy <- temp %>% mutate(Status = ML_set$Status, Complicated_Status = ML_set$Complicated_Status)
# recreate for full data
temp <- all_set %>% dplyr::select(-filter, -Status, -Complicated_Status)
temp <- dummy.data.frame(temp, sep='_')
all_dummy <- temp %>% mutate(Status = all_set$Status, Complicated_Status = all_set$Complicated_Status)
```

# Identify highly correlated metrics
```{r}
correlated <- cor(all_dummy %>% select(-Status, -Complicated_Status))
indices <- findCorrelation(correlated)
colnames(all_dummy %>% select(-Status, -Complicated_Status))[indices]
```

## Train and Test sets
Split in training / testing

```{r}
set.seed(115470)
train_set <- ML_set_dummy %>% 
  group_by(Status, Complicated_Status) %>% 
 # filter(!Complicated_Status=='Comp_Het') %>% # remove comp hets for now
  sample_frac(0.5) %>% ungroup()
# remove no variance columns
#var_columns <- colnames(train_set)[apply(train_set, MARGIN = 2, function(x) var(x) > 0)]
#var_columns <- var_columns[!is.na(var_columns)]
#train_set <- train_set[,c(var_columns, 'Status', 'Complicated_Status')]


test_set <- ML_set_dummy %>% 
  filter(!variant_id %in% train_set$variant_id)
# remove no variance columns
#var_columns <- colnames(test_set)[apply(test_set, MARGIN = 2, function(x) var(x) > 0)]
#var_columns <- var_columns[!is.na(var_columns)]
#test_set <- test_set[,c(var_columns, 'Status', 'Complicated_Status')]
```

## Recipe
skip for now
```{r, eval=FALSE}
library(caret)
train_setR <- train_set %>% dplyr::select(-Complicated_Status, -variant_id)
rec_obj <- recipe(Status ~ ., data = train_setR)
```



## Modeling

```{r}
library(caret)
library(mlbench)
library(parallel)
library(doParallel)
library(MLmetrics)
#library(plotROC)


fitControl <- trainControl(## 5-fold CV
 # method = "cv",
 # number = 5,
  classProbs=T,
  savePredictions = T,
  allowParallel = T,
  summaryFunction = twoClassSummary)
cluster <- makeCluster(detectCores() - 1) # leave some cores for me!
registerDoParallel(cluster)

set.seed(825)
columns <- c('type_indel','type_snp','num_hom_ref','num_het','num_hom_alt','aaf','is_lof','is_splicing','impact_frameshift_variant','impact_inframe_deletion','impact_inframe_insertion','impact_intron_variant','impact_missense_variant','impact_splice_acceptor_variant','impact_splice_donor_variant','impact_splice_region_variant','impact_start_lost','impact_stop_gained','impact_stop_lost','impact_synonymous_variant','aaf_1kg_all_float','aaf_esp_all','af_exac_all','pli','pnull','precessive','phylop_100way','cadd_phred', 'polyphen_score', 'impact_severity_HIGH', 'impact_severity_LOW', 'impact_severity_MED', 'polyphen_score', 'sift_score', 'fitcons_float','lof_z','mis_z','syn_z','gerp_elements','grantham')
rfFit <- caret::train(x=train_set %>% 
                        dplyr::select_(.dots=columns) %>% 
                        data.frame(), 
                      y = train_set$Status, 
                      #preProcess = c("scale", "center"),
                      method = "rf", 
                      trControl = fitControl)

bglmFit <- caret::train(x=train_set %>% 
                        dplyr::select_(.dots=columns) %>% 
                        data.frame(), 
                      y = train_set$Status, 
                      #preProcess = c("scale", "center"),
                      method = "bayesglm",
                      trControl = fitControl)
fdaFit <- caret::train(x=train_set %>% 
                        dplyr::select_(.dots=columns) %>% 
                        data.frame(), 
                      y = train_set$Status, 
                      #preProcess = c("scale", "center"),
                      method = "fda", 
                      trControl = fitControl)
xgbTreeFit <-caret::train(x=train_set %>% 
                        dplyr::select_(.dots=columns) %>% 
                        data.frame(), 
                      y = train_set$Status, 
                     # preProcess = c("scale", "center"),
                      method = "xgbTree", 
                      trControl = fitControl)
c5.0TreeFit <- caret::train(x=train_set %>% 
                        dplyr::select_(.dots=columns) %>% 
                        data.frame(), 
                      y = train_set$Status, 
                      #preProcess = c("scale", "center"),
                      method = "C5.0Tree",
                      trControl = fitControl)
stepLDAFit <- caret::train(x=train_set %>% 
                        dplyr::select_(.dots=columns) %>% 
                        data.frame(), 
                      y = train_set$Status, 
                      #preProcess = c("scale", "center"),
                      method = "stepLDA", 
                      trControl = fitControl)
# cadd only
caddFit <- caret::train(x=train_set %>% 
                        dplyr::select(cadd_phred) %>% 
                        data.frame(), 
                      y = train_set$Status, 
                      #preProcess = c("scale", "center"),
                      method = "glm", 
                      trControl = fitControl)

#stopCluster(cluster)
#registerDoSEQ() # forces R down to single thread - can't do parallel run after this?

my_models <- list(rfFit,bglmFit,xgbTreeFit,c5.0TreeFit,stepLDAFit,caddFit)
names(my_models) <- c('rfFit','bglmFit','xgbTreeFit','c5.0TreeFit','stepLDAFit','caddFit')
```

# ConfusionMatrix your model(s)
```{r}
# fuller_data is all variants, minus the variants in the test_set
fuller_data = all_dummy %>% filter(!variant_id %in% test_set$variant_id)

cm_maker <- function(model, data, cutoff=0.5) {
  new_predictions <- predict(model, data, type='prob') %>%
    mutate(Answers = data$Status, Prediction = case_when(Pathogenic > cutoff ~ 'Pathogenic', TRUE ~ 'NotPathogenic'))
  confusionMatrix(data = new_predictions$Prediction, reference = new_predictions$Answers, mode='prec_recall')
}
#example
cm_maker(bglmFit, fuller_data, cutoff=0.5)
```

# AUC ROC
```{r}
library(PRROC)

# precision recall AUC
aucroc_maker <- function(model, data, cutoff=0.5) {
  new_predictions <- predict(model, data, type = 'prob') %>%
    mutate(Answers = data$Status, Prediction = case_when(Pathogenic > cutoff ~ 'Pathogenic', TRUE ~ 'NotPathogenic'))
  pr.curve(scores.class0 = new_predictions %>% filter(Answers=='Pathogenic') %>% pull(Pathogenic),
           scores.class1 = new_predictions %>% filter(Answers=='NotPathogenic') %>% pull(Pathogenic),
           curve = T)
}

# ROC AUC
roc_maker <- function(model, data, cutoff=0.5) {
  new_predictions <- predict(model, data, type = 'prob') %>%
    mutate(Answers = data$Status, Prediction = case_when(Pathogenic > cutoff ~ 'Pathogenic', TRUE ~ 'NotPathogenic'))
  roc.curve(scores.class0 = new_predictions %>% filter(Answers=='Pathogenic') %>% pull(Pathogenic),
           scores.class1 = new_predictions %>% filter(Answers=='NotPathogenic') %>% pull(Pathogenic),
           curve = T)
}

aucroc_data <- data.frame()
for (i in names(my_models)){
  print(my_models[[i]]$method)
  out <- aucroc_maker(my_models[[i]], fuller_data)
  out <- out$curve[,1:2] %>% data.frame()
  colnames(out) <- c('Recall','Precision')
  out$model <- i
  aucroc_data <- rbind(aucroc_data, out)
}

pr <- aucroc_data %>% ggplot(aes(x=Recall, y=Precision, colour=model)) + geom_line() + theme_minimal() + ggsci::scale_color_aaas()


roc_data <- data.frame()
for (i in names(my_models)){
  print(my_models[[i]]$method)
  out <- roc_maker(my_models[[i]], fuller_data)
  out <- out$curve[,1:2] %>% data.frame()
  colnames(out) <- c('FPR','Sensitivity')
  out$model <- i
  roc_data <- rbind(roc_data, out)
}

roc <- roc_data %>% ggplot(aes(x=FPR, y=Sensitivity, colour=model)) + geom_line() + theme_minimal() + ggsci::scale_color_aaas()
library(cowplot)
cowplot::plot_grid(roc, pr)


```
