---
title: "RF Model"
output: html_notebook
---




# Load data from stats.Rmd
```{r}
load('/Volumes/Arges_NFS/PROJECTS/mcgaughey/EGA_EGAD00001002656_NGS_reanalyze/clean_data/gemini_out_variants_het_and_hom.Rdata')
```

# Set up data for modeling
```{r}
library(tidyverse)
all_PATH <- all %>% 
  select(variant_id, filter, type, call_rate, num_hom_ref, num_het, num_hom_alt, aaf, is_exonic, is_coding, is_lof, is_splicing, impact, impact_so, impact_severity, polyphen_score, sift_score, aaf_1kg_afr_float:an_exac_sas, max_aaf_all, pli, pnull, precessive, phylop_100way, canonical, cadd_phred, Genotype, Status, Complicated_Status) %>% 
  filter(Status=='Pathogenic') %>% 
  unique()
all_NOT_PATH <- all %>% 
  select(variant_id, filter, type, call_rate, num_hom_ref, num_het, num_hom_alt, aaf, is_exonic, is_coding, is_lof, is_splicing, impact, impact_so, impact_severity, polyphen_score, sift_score, aaf_1kg_afr_float:an_exac_sas, max_aaf_all, pli, pnull, precessive, phylop_100way, canonical, cadd_phred, Genotype, Status, Complicated_Status) %>% 
  filter(Status=='NotPathogenic') %>% 
  unique()
all_set <- rbind(all_PATH, all_NOT_PATH)
process <- function(df) {
  #cadd phred has NA
  df[is.na(df)] <- 0
  # convert chr to factor
  df <- df %>%  mutate_each_( funs(as.factor(.)), names( .[,sapply(., is.character)] ))
  # Fix numeric that are fctr
  df$polyphen_score <- as.numeric(df$polyphen_score)
  df$sift_score <- as.numeric(df$sift_score)
  df$adj_exp_lof <- as.numeric(df$adj_exp_lof)
  df$adj_exp_mis <- as.numeric(df$adj_exp_mis)
  df$adj_exp_syn <- as.numeric(df$adj_exp_syn)
  df$pli <- as.numeric(df$pli)
  df$pnull <- as.numeric(df$pnull)
  df$precessive <- as.numeric(df$precessive)
  df$phylop_100way <- as.numeric(df$phylop_100way)
  df$Status <- factor(df$Status, levels=c('Pathogenic','NotPathogenic'))
  df
}

all_set <- process(all_set)

# multiply path by 10x with duplications
#all_PATH_10X <- bind_rows(replicate(10, all_PATH, simplify = FALSE))
all_PATH_10X <- all_set %>% filter(Status=='Pathogenic')
# cut down pathogenic to 5x of path
set.seed(115470)
all_NOT_PATH_CUT <- all_set %>% filter(Status=='NotPathogenic') %>% sample_n(nrow(all_PATH_10X) * 5)

ML_set <- rbind(all_PATH_10X, all_NOT_PATH_CUT)



```

## One hot encoding (Dummy variables)
Turn categorical factors into variables

Many ML algorithms can't handle factor

You can feed this into the `Train and Test sets` section below if you are using something like a GLM or NN. 

Skip for now
```{r, eval = F}
library(caret)
temp <- ML_set %>% dplyr::select(-filter, -Status, -Complicated_Status)
dmy <- dummyVars(" ~ .", data = temp, fullRank=T)
ML_set_dummy <- data.frame(predict(dmy, newdata = temp))
ML_set_dummy <- ML_set_dummy %>% mutate(Status = ML_set$Status, Complicated_Status = ML_set$Complicated_Status)
# recreate for full data
temp <- rbind(all_PATH, all_NOT_PATH) %>% dplyr::select(-filter, -Status, -Complicated_Status)
dmy <- dummyVars(" ~ .", data = temp, fullRank=T)
all_dummy <- data.frame(predict(dmy, newdata = temp))
all_dummy <- ML_set_dummy %>% mutate(Status = ML_set$Status, Complicated_Status = ML_set$Complicated_Status)
```

## Train and Test sets
Split in training / testing

```{r}
set.seed(115470)
train_set <- ML_set %>% 
  group_by(Status, Complicated_Status) %>% 
 # filter(!Complicated_Status=='Comp_Het') %>% # remove comp hets for now
  sample_frac(0.5) %>% ungroup()
# remove no variance columns
#var_columns <- colnames(train_set)[apply(train_set, MARGIN = 2, function(x) var(x) > 0)]
#var_columns <- var_columns[!is.na(var_columns)]
#train_set <- train_set[,c(var_columns, 'Status', 'Complicated_Status')]


test_set <- ML_set %>% 
  filter(!variant_id %in% train_set$variant_id)
# remove no variance columns
#var_columns <- colnames(test_set)[apply(test_set, MARGIN = 2, function(x) var(x) > 0)]
#var_columns <- var_columns[!is.na(var_columns)]
#test_set <- test_set[,c(var_columns, 'Status', 'Complicated_Status')]
```

## Recipe
skip for now
```{r, eval=FALSE}
library(caret)
train_setR <- train_set %>% dplyr::select(-Complicated_Status, -variant_id)
rec_obj <- recipe(Status ~ ., data = train_setR)
```



## Modeling

```{r}
library(caret)
library(mlbench)
library(parallel)
library(doParallel)
library(MLmetrics)
#library(plotROC)


fitControl <- trainControl(## 5-fold CV
 # method = "cv",
 # number = 5,
  classProbs=T,
  savePredictions = T,
  allowParallel = T)
cluster <- makeCluster(detectCores() - 1) # leave some cores for me!
registerDoParallel(cluster)

set.seed(825)
rfFit <- caret::train(x=train_set %>% dplyr::select(type, num_hom_ref, num_het, num_hom_alt, aaf, is_exonic, is_coding, is_lof, is_splicing, impact, polyphen_score, sift_score, max_aaf_all, pli, pnull, precessive, phylop_100way, cadd_phred) %>% data.frame(), 
                      y = train_set$Status, 
                      #preProcess = c("scale", "center"),
                      method = "rf", 
                      trControl = fitControl)
bglmFit <- caret::train(x=train_set %>% dplyr::select(type, num_hom_ref, num_het, num_hom_alt, aaf, is_exonic, is_coding, is_lof, is_splicing, impact, polyphen_score, sift_score, max_aaf_all, pli, pnull, precessive, phylop_100way, cadd_phred) %>% data.frame(), 
                      y = train_set$Status, 
                      #preProcess = c("scale", "center"),
                      method = "bayesglm", 
                      trControl = fitControl)
xgbTreeFit <- caret::train(x=train_set %>% dplyr::select(type, num_hom_ref, num_het, num_hom_alt, aaf, is_exonic, is_coding, is_lof, is_splicing, impact, polyphen_score, sift_score, max_aaf_all, pli, pnull, precessive, phylop_100way, cadd_phred)  %>% data.frame(), 
                      y = train_set$Status, 
                     # preProcess = c("scale", "center"),
                      method = "xgbTree", 
                      trControl = fitControl)
svmFit <- caret::train(x=train_set %>% dplyr::select(type, num_hom_ref, num_het, num_hom_alt, aaf, is_exonic, is_coding, is_lof, is_splicing, impact, polyphen_score, sift_score, max_aaf_all, pli, pnull, precessive, phylop_100way, cadd_phred)  %>% data.frame(), 
                      y = train_set$Status, 
                      #preProcess = c("scale", "center"),
                      method = "svmLinear", 
                      trControl = fitControl)



stopCluster(cluster)
registerDoSEQ() # forces R down to single thread - can't do parallel run after this?
```

# ConfusionMatrix your model(s)
```{r}
cm_maker <- function(model, data, cutoff=0.5) {
  new_predictions <- predict(model, data, type='prob') %>%
    mutate(Answers = data$Status, Prediction = case_when(Pathogenic > cutoff ~ 'Pathogenic', TRUE ~ 'NotPathogenic'))
  confusionMatrix(data = new_predictions$Prediction, reference = new_predictions$Answers, mode='prec_recall')
}
# change cutoff
confusionMatrix(data = rfFit$pred %>% mutate(pred2 = case_when(Pathogenic > 0.2 ~ 'Pathogenic', TRUE ~ 'NotPathogenic')) %>% pull(pred2), reference = rfFit$pred$obs)

randomForest::varImpPlot(rfFit_tissue$finalModel)
selectedIndices <- rfFit$pred$mtry == rfFit_tissue$bestTune[1,] #26

g <- ggplot(rfFit$pred[selectedIndices, ], aes(m=Pathogenic, d=factor(obs, levels=c('Pathogenic','NotPathogenic')))) + 
  geom_roc(n.cuts=0) + 
  coord_equal() +
  style_roc()

g + annotate("text", x=0.75, y=0.25, label=paste("AUC =", round((calc_auc(g))$AUC, 4)))
```